Learning Reusable Components for Neurally--Guided Bayesian Program Learning

### 2018-05-12 nips_2018.tex
DSL fragment example

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\newcommand{\theHalgorithm}{\arabic{algorithm}}

\newcommand{\system}{\textsc{ECC}~}
\newcommand{\systemEnding}{\textsc{ECC}}
\newcommand{\lowerBound}{\mathscr{L}}
\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
\newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}

@@

The above equations summarize the problem from the point of view of an ideal Bayesian learner.
However, Eq.~\ref{intractableObjectives}
is wildly intractable because evaluating $J(\mathcal{D},\theta)$ involves
summing over the  infinite set of all programs.
In practice we will only ever be able to sum over a finite set of programs.
So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
\\\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
is a finite set of programs where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.

@@

Searching for programs is hard because
of the large combinatorial search space. We ease this difficulty by training a neural recognition model
during the compilation phase:

@@

\subsection{Compilation: Learning a Recognition Model}\label{recognitionSection}

The purpose of the recognition model is to accelerate the search for
programs.  It does this by learning to predict programs which are
probable under $(\mathcal{D},\theta)$ while also assigning high
likelihood for a task according to $\probability[x|p]$.  Concretely,
the recognition model $q$ is a neural network that predicts, for each
task $x\in X$, a weight vector $q(x) = \theta^{(x)}\in
\mathbb{R}^{|\mathcal{D}| + 1}$.  Together with the DSL, this defines
a distribution over programs, $\probability[p|\mathcal{D},\theta =
  q(x)]$.  We abbreviate this distribution as $q(p|x)$.  The crucial
aspect of this framing is that the neural network leverages the
structure of the learned DSL, so it is \emph{not} responsible for
generating programs wholesale.  We share this aspect with
DeepCoder~\cite{balog2016deepcoder}.

We want a recognition model that closely approximates the true posteriors over programs. We formulate this as minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
or equivalently maximizing
\begin{equation*}
  \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
\end{equation*}
where the expectation is taken over tasks. One could take this expectation
over the observed empirical distribution of tasks,
like how an autoencoder is trained~\cite{hinton2006reducing}; or, one could take this expectation over samples from the generative model, like how a Helmholtz machine is trained~\cite{dayan1995helmholtz}.
We found it useful to maximize both an autoencoder-style objective $\mathcal{L}_{\text{AE}}$ and a Helmholtz-style objective $\mathcal{L}_{\text{HM}}$, giving the  objective for a recognition model, $\mathcal{L}_{\text{RM}} = \mathcal{L}_\text{AE} + \mathcal{L}_\text{HM}$:
\begin{align}
\mathcal{L}_{\text{HM}} = \expect_{p\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right],\text{ $p$ evaluates to $x$}\nonumber\qquad
\mathcal{L}_{\text{AE}} = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]
\end{align}
The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
all of our experiments train \system on only a few hundred tasks, which is too little for
a high-capacity neural network $q$.
Once we bootstrap a $(\mathcal{D},\theta)$,
we can draw unlimited samples from $(\mathcal{D},\theta)$
and train $q$ on those samples.

@@

\begin{figure*}
\centering  \begin{tabular}{lll}
    \toprule
    Domain&Example programs in frontiers&Example proposed subexpression\\\midrule
    Circuits&
    \begin{tabular}{l}
      \code{(lambda x (nand x x))}\\
      \code{(lambda x (lambda y (nand x (nand y y))}
    \end{tabular}&
    \code{(nand z z)}\\\midrule
    Strings&
    \begin{tabular}{l}
      \code{(lambda s (+ ',' (index 0 (split ',' s))))}\\
      \code{(lambda s (index 0 (split ' ' s)))}
    \end{tabular}&
    \code{(index 0 (split z s))}\\\bottomrule
  \end{tabular}
  \caption{The DSL induction algorithm proposes subexpressions of programs to add to the DSL.
    These subexpressions are taken from programs in the frontiers (middle column), and can introduce new variables (\code{z} and \code{s} in the right column).} 
  \label{fragmentExample}
\end{figure*}

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{stmaryrd}

\newcommand{\theHalgorithm}{\arabic{algorithm}}
\newcommand{\sem}[1]{\llbracket #1 \rrbracket}
\newcommand{\system}{\textsc{ECC}~}
\newcommand{\systemEnding}{\textsc{ECC}}
\newcommand{\lowerBound}{\mathscr{L}}
\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
\newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}

@@

The above equations summarize the problem from the point of view of an ideal Bayesian learner.
However, Eq.~\ref{intractableObjectives}
is wildly intractable because evaluating $J(\mathcal{D},\theta)$ involves
summing over the  infinite set of all programs.
In practice we will only ever be able to sum over a finite set of programs.
So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
\\\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
is a finite set of programs s.t. $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.

@@

Searching for programs is hard because
of the large combinatorial search space. We ease this difficulty by training a neural recognition model, $q(\cdot |\cdot )$,
during the compilation phase: $q$ is trained to approximate the
posterior over programs, $q(p|x)\propto\probability[x,p|\mathcal{D},\theta}]$,
  thus amortizing the cost of finding programs with high posterior probability.

@@

\subsection{Compilation: Learning a Neural Recognition Model}\label{recognitionSection}

The purpose of the recognition model is to accelerate the search for
programs.  It does this by learning to predict programs which are
probable under $(\mathcal{D},\theta)$ while also assigning high
likelihood for a task according to $\probability[x|p]$.  Concretely,
the recognition model $q$ is a neural network that predicts, for each
task $x\in X$, a weight vector $q(x) = \theta^{(x)}\in
\mathbb{R}^{|\mathcal{D}| + 1}$.  Together with the DSL, this defines
a distribution over programs, $\probability[p|\mathcal{D},\theta =
  q(x)]$.  We abbreviate this distribution as $q(p|x)$.  The crucial
aspect of this framing is that the neural network leverages the
structure of the learned DSL, so it is \emph{not} responsible for
generating programs wholesale.  We share this aspect with
DeepCoder~\cite{balog2016deepcoder} and~\cite{menon2013machine}.

We want a recognition model that closely approximates the true posteriors over programs. We formulate this as minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
or equivalently maximizing
\begin{equation*}
  \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
\end{equation*}
where the expectation is taken over tasks. One could take this expectation
over the observed empirical distribution of tasks,
like how an autoencoder is trained~\cite{hinton2006reducing}; or, one could take this expectation over samples from the generative model, like how a Helmholtz machine is trained~\cite{dayan1995helmholtz}.
We found it useful to maximize both an autoencoder-style objective $\mathcal{L}_{\text{AE}}$ and a Helmholtz-style objective $\mathcal{L}_{\text{HM}}$, giving the  objective for a recognition model, $\mathcal{L}_{\text{RM}} = \mathcal{L}_\text{AE} + \mathcal{L}_\text{HM}$:
\begin{align}\nonumber
\mathcal{L}_{\text{HM}} = \expect_{p\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right]\text{ s.t. $x = \sem{p}$}\quad
\mathcal{L}_{\text{AE}} = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]
\end{align}
The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
all of our experiments train \system on only a few hundred tasks, which is too little for
a high-capacity neural network $q$.
Once we bootstrap a $(\mathcal{D},\theta)$,
we can draw unlimited samples from $(\mathcal{D},\theta)$
and train $q$ on those samples.

@@

\begin{figure*}
\centering  \begin{tabular}{ll}
    \toprule
    Example programs in frontiers&Proposed subexpression\\\midrule
    \begin{tabular}{l}
      \code{(lambda (a b) (fold b (cons "," a) (lambda (x z) (cons x z))))}\\
      \code{(lambda (a b) (fold a b (lambda (x z) (cons x z))))}
    \end{tabular}&
    \code{(fold a b (lambda (x z) (cons x z)))}\\\bottomrule
  \end{tabular}
  \caption{The DSL induction algorithm proposes subexpressions of programs to add to the DSL.
    These subexpressions are taken from programs in the frontiers (left column), and can introduce new variables (right column: \code{a} and \code{b}). Here, the proposed subexpression appends two lists.} 
  \label{fragmentExample}
\end{figure*}
